Forecasting the 2024 U.S. Presidential Election: Polling Trends and Dynamics Based on National General Election Polls

### 2024-11-04 paper/paper.qmd
Updated Paper

library(tidyverse)
library(usmap)
library(dplyr)
library(here)
library(ggplot2)
library(janitor)
library(knitr)
library(flextable)
library(arrow)
analysis_data <- read_parquet(here("data/02-analysis_data/analysis_data.parquet"))
```

@@

```{r visualize-sample-size}
# Ensure 'analysis_data' is already loaded at this point
ggplot(analysis_data, aes(x = sample_size)) +
  geom_histogram(binwidth = 100, fill = "skyblue", color = "black") +
  labs(title = "Distribution of Sample Sizes in Polls", 
       x = "Sample Size", 
       y = "Frequency") +
  theme_minimal()

@@

```{r}
# Visualization: Candidate Support by State
filtered_data <- analysis_data %>% filter(state != "National")
ggplot(filtered_data, aes(x = state, y = pct, fill = candidate_name)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Candidate Support by State", 
       x = "State", 
       y = "Percentage Support") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), 
        legend.position = "top")
```

@@

>>>>

library(tidyverse)
library(usmap)
library(dplyr)
library(here)
library(ggplot2)
library(janitor)
library(knitr)
library(flextable)
library(arrow)

#Reads the data and store it into analysis_data 
analysis_data <- read_parquet(here("data/02-analysis_data/analysis_data.parquet"))
```

@@

```{r visualize-sample-size}
#| warning: false
#| message: false
#| echo: false
# Ensure 'analysis_data' is already loaded at this point
ggplot(analysis_data, aes(x = sample_size)) +
  geom_histogram(binwidth = 100, fill = "skyblue", color = "black") +
  labs(title = "Distribution of Sample Sizes in Polls", 
       x = "Sample Size", 
       y = "Frequency") +
  theme_minimal()

@@

```{r}
#| warning: false
#| message: false
#| echo: false
# Visualization: Candidate Support by State
filtered_data <- analysis_data %>% filter(state != "National")
ggplot(filtered_data, aes(x = state, y = pct, fill = candidate_name)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Candidate Support by State", 
       x = "State", 
       y = "Percentage Support") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), 
        legend.position = "top")

@@

## Purpose

This paper analyzes and predicts the 2024 U.S. presidential election by fitting a regression model based on polling data downloaded from FiveThirtyEight (@fivethirtyeight). Our analysis focuses on the leading candidates, examining key variables such as the pollster, the date the poll was conducted, sample size, and the percentage of respondents supporting each candidate.

Additionally, this analysis aimed to estimate the 2024 U.S. presidential election electoral college outcome by propagating uncertainty through polling data. Using Monte Carlo simulations, we captured the variability in candidate support across states and produced a distribution of possible electoral outcomes. This integration enabled us to estimate electoral votes for each candidate by state, taking into account each state's unique polling landscape.

## Limitations

While Monte Carlo simulations provide a robust framework for propagating uncertainty, the reliability of our estimates is reliant upon the accuracy and representatives of polling data.

As our estimates are contingent upon the accuracy and representativeness of polling data, we cannot guarantee absolute precision in our projections. Nate Silver (2024) highlights a particular challenge in polling known as "herding," where polling organizations may unintentionally align their results to avoid being statistical outliers (@NYMagHerding2024).This alignment can distort the true range of our estimates. Since our projections suggest that Trump would win, it’s important to address the role of standard deviations in our estimates and consider the possibility that Harris’s support may be underestimated, similar to how Trump’s support was in 2020 (@FTArticle2024). The idea of Harris being underestimated, much like Trump was in 2020, hinges on the possibility that polling errors or biases could be systematically missing certain groups of voters who support her. In 2020, Trump’s support was underestimated in some key swing states due to factors such as nonresponse bias, where Trump-leaning voters were less likely to participate in polls, and demographic weighting errors.
